A subjective Bayesian framework for synthesizing deep uncertainties in climate risk management

### 2022-09-20 papers/earths-future/reviewer_response_01.tex

\usepackage{siunitx}
\usepackage{apacite}
\usepackage{physics}

\usepackage[colorinlistoftodos]{todonotes}
\usepackage{color}
\newcommand{\jdg}[1]{\todo[color=red!30]{#1}}
\newcommand{\all}[1]{\todo[color=blue!30]{#1}}

% better lists
\usepackage{enumitem}
\setlist{nosep}

% some commands
\usepackage{xspace}
\makeatletter
\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}
\def\eg{\emph{e.g}\onedot} \def\Eg{\emph{E.g}\onedot}
\def\ie{\emph{i.e}\onedot} \def\Ie{\emph{I.e}\onedot}
\def\etc{\emph{etc}\onedot} \def\vs{\emph{vs}\onedot}
\newcommand{\usd}[1]{\SI{#1}[\$]{}}

% ACRONYMS
\usepackage[acronym, nopostdot, nonumberlist, shortcuts, numberedsection, nogroupskip,]{glossaries}
\newacronym{bfe}{BFE}{base flood elevation}
\newacronym{cmip}{CMIP}{the Coupled Model Intercomparison Project}
\newacronym{dmdu}{DMDU}{decision making under deep uncertainty}
\newacronym{fema}{FEMA}{the Federal Emergency Management Agency}
\newacronym{gev}{GEV}{generalized extreme value}
\newacronym{hazus}{HAZUS}{Hazard U.S.}
\newacronym{iid}{IID}{independent and identically distributed}
\newacronym{ipcc}{IPCC}{International Panel on Climate Change}
\newacronym{msl}{MSL}{mean relative sea level}
\newacronym{noaa}{NOAA}{the National Oceanic and Atmospheric Administration}
\newacronym{pdf}{PDF}{probability density function}
\newacronym{rcp}{RCP}{representative concentration pathway}
\newacronym{rdm}{RDM}{robust decision making}
\newacronym{slr}{SLR}{sea level rise}
\newacronym{ssp}{SSP}{shared socio-economic pathway}
\newacronym[]{usace}{USACE}{United States Army Corps of Engineers}
\newacronym[]{usgs}{USGS}{United States Geological Survey}
\newacronym[\glslongpluralkey={states of the world}]{sow}{SOW}{state of the world}

@@

Referee 2 encouraged us to (1) better contextualize our methodological approach relative to previous works, (2) clarify and improve the mathematical formulation of our approach, and (3) to better describe the usefulness of our method.
The specific

In the remainder of this document we address the specific comments made by the two referees.

\section{Referee \#1}

\RC{
    This work introduces a Bayesian framework to synthesize and characterize deep uncertainties to merge standard exploratory modeling techniques with more traditional decision support tools that require statistics like the expected value.
    Such statistics, which require weighting sampled states of the world, are problematic in the presence of deep uncertainty.
    The authors argue that many decision making frameworks commonly used for robust decision-making problems actually make such a weighting, but that is not transparent.
    Decisions about the sampling range, joint sampling distribution, and statistics (such as domain criterion) can have consequences for the robustness of the options presented to stakeholders, but this may not be clear.
    The proposed framework allows for a synthesis across deep uncertainties using subjective Bayesian priors.
    The method is demonstrated for a very simple house-raising example including both sea level rise and a stationary storm surge risk.

    Overall, I found the work well written and informative, and I think it is very promising.
    The framework's novelty arises from the way it stitches standard methods from Bayesian statistics and DMDU.
    This is not a negative statement, as I think by employing common-sense methods in a novel way the authors' framework is more likely to gain traction.
    An important contribution is the introduction, which nicely lays out the current state of practice, emerging methods (primarily from the DMDU community), and then research gaps.
    I appreciated the author's care in discussing the difficulty in adopting non-stationary frequency methods for extremes, with key citations to important works by Montanari, Seranaldi, and Bulletin 17C's introduction.
}

@@

\AR{
    Thank you for your comments.
    This is an excellent summary of the aim and scope of this paper.
}

\RC{
    The use of the word model to mean different things. Throughout the manuscript, and particularly in Figure 1, the authors utilize the term `model' to refer to both the combination of RCP-SLR model/assumption and the systems model.
    I found this confusing at multiple places.
    One suggestion is to call the 'models' in box d of Figure 1 'scenarios' or something like that.
    So the scenario is the deep uncertainty (RCP-SLR model) and then the possible futures in box b of Figure 1 are samples drawn from the pdf that's conditioned on the scenario in Box d.
    I think this new terminology would be particularly important if one were to have more than one systems model in box c.
}

@@

\RC{
    Figure 8: The y-axis labels on B-D should be more descriptive.
}

\AR{
    Yes, can do.
}

\section{Referee \#2}

\RC{
    This manuscript addresses an important topic: the role of probabilistic methods in decision-making under deep uncertainty.
    The authors point out that scenario-based DMDU methods implicitly put probabilities on scenarios but do not make these assumptions transparent.
    They address this by developing a method to calculate the weights implicitly applied to each deeply uncertain scenario.
    I applaud this effort -- the tension between scenario-based methods that use optimization and probabilistic methods is an important and understudied one.
    I also see a lot of value in how the authors have framed the problem and described the pros and cons of scenario-based DMDU methods and probabilistic methods for CBA in the context of infrastructure design.
    The simple case study of raising the elevation of a house under flood risk affected by deeply uncertain sea level rise is appropriate for demonstrating the method and is technically sound.

    However, I have three major concerns that prevent me from recommending the manuscript for publication in its current form.
    I defer to the editor on whether, assuming all these concerns are addressed in revision, the contribution warrants publication in Earth's Future.
}

@@

\AR{
    Thanks for the insightful comments.
}

\RC{
    First, the paper is not well grounded in previous literature at the intersection of Bayesian /probabilistic methods and DMDU.
    This has led the authors to make overly broad claims about their contribution that make it difficult to discern what they are adding beyond existing methods.
    It is certainly not true that ``In this paper we offer a first conceptual step towards bridging the fields of DMDU and Bayesian model building'' as claimed in line 183.
    Here are some papers that integrate Bayesian/probabilistic methods into DMDU contexts:
    \citeA{reis_distribution:2020},
    \citeA{taner_probabilistic:2019},
    \citeA{hui_adaptive:2018},
    \citeA{fletcher_learning:2019}, and
    \citeA{fletcher_groundwater:2019}.
    The first three are especially similar to the authors' contributions; the authors need to make a compelling argument for what their approach adds beyond this existing literature.
}

\AR{
    Thanks very much for this helpful comment.
    This comment underscores a need for improved clarity in the introduction to the manuscript; as noted, there are numerous papers that use probabilistic tools, including Bayesian methods, in DMDU contexts.
    We have clarified that there is a long literature on the relative merits of probabilistic vs non-probabilistic information in DMDU contexts, as noted in \citeA{taner_probabilistic:2019}, but that this debate has often centered on a top-down, technocratic, objectivist philosophy of probability in which there is explicitly or implicitly assumed to be a true probability distribution.
    Our contribution is to outline how a subjective Bayesian philosophical approach can address many of the concerns raised by critics of probabilistic approaches while providing a principled and transparent method for reasoning about uncertainty.

    We can add a paragraph to the introduction describing this intersection, possibly with a few additional papers.
    \citeA{hui_adaptive:2018} is a nice example of the use of SDP applied to a levee heightening problem.
    There is Bayesian updating through the dynamic programming, but deep uncertainty is addressed only by considering several scenarios of hydrological conditions.
    \citeA{taner_probabilistic:2019} is conceptually quite similar to our approach.
    This paper focuses on joint distributions of uncertain parameters rather than how different subjective beliefs about these parameters affect the decision.
    This paper also focuses on how probabilistic versus exploratory frameworks lead to different robustness values, whereas here we focus more generally on mapping trade-offs.
    \citeA{reis_distribution:2020} shows that scenario discovery and exploratory modeling results are sensitive to parameterization and choice of distribution.
    \citeA{fletcher_learning:2019} and \citeA{fletcher_groundwater:2019} apply SDP (which can be interpreted as having an implicit Bayesian updating step) but shallow uncertainties are considered.
}

\begin{quote}
    Add a sentence to the intro and link it here
\end{quote}

@@

\RC{
    My second key concern is that the mathematical formulation is not clear and complete, and it is not always clear how the sample application aligns with the general framework.
    Given that the authors suggest the mathematical formulation is a key part of their contribution, this is important.
    Specifically:
    The method is framed both in the introduction (lines 157-182) and methods section (lines 240-250) around Bayesian model selection.
    However, it is not clear if or how the authors are performing Bayesian model selection.
    In figure 1e and line 247, the authors provide a conditional probability of $u$ given $x_i$ and $M_k$, and state this is the result of IID draws from $M_k$.
    This is similar to the $p(y|M_k)$ term in the classical BMA framework from \citeA{raftery_bma:2005} equation 1.
    Does this imply that there is a Bayesian posterior for model weights developed?
    If so, what is it?
    What data is used for the updating and what is the likelihood function? It is unclear if or how Bayes' is used here.
}

\AR{
    This is a great question and one we need to clarify.
    For reference, equation 1 of \citeA{raftery_bma:2005} is
    $$p(y) = \frac{1}{K} \sum_{k=1}^K p(y|M_k) p(M_k | y^T),$$
    ``where $p( y | M_k)$ is the forecast PDF based on model $M_k$ alone, and $p(M_k | y^T)$ is the posterior probability of model $M_k$ being correct given the training data, and reflects how well model $M_k$ fits the training data.''

    This comment is closely related to the previous comment in that we draw inspiration from the literature on Bayesian model selection but do not explicitly perform it.
    Our approach is conceptually similar to BMA as noted (and even close to stacking; see \cite{Yao:2018bu}).
    There are two key differences.
    First, these methods focus on estimating the weights assigned to each model from data; in our context we have no data.
    Future work could consider adaptive decision making (by revisiting the decision at multiple time steps, \ie stochastic control) in which case there would be data available in the future that could be used to estimate these weights as in several of the papers referenced in the previous comment.
    Second, these methods assign weights to each model, whereas we assign weights to each SOW.
    This is motivated by the fact that the SOWs in our applications, and in many climate adaptation applications, are computationally expensive to generate.
}

\begin{quote}
    Need to think about how we want to address this in the text
\end{quote}

\RC{
    Shortly later in the paragraph under section 2.3, the authors take an expected value of $f(s)$ and state that it ``can be readily estimated as $\frac{1}{N} \sum_{j=1}^N f(\bf{s}_j)$.''
    I believe this is inconsistent with the interpretation of the SOW being samples from an underlying distribution.
    We know that:
    $$\mathbb{E}\qty[f(s)]=\sum_{s \in \Omega}f(s)P_s(s)$$
    where $P_s$ is the probability mass function of $s$.
    This is equivalent to the authors' equation above only when assuming that $P_s(s)$ is a discrete uniform RV where each SOW gets equal probability.
    However, this seems to be at odds with the authors interpretation that the SOW are IID draws from some underlying distribution of models.
    If that is the case, the probability mass or density function shouldn't simply put probability on the available SOW, as these are merely samples from an underlying distribution.
    It conflates the sample with the population.
    If, for example, the SOW were believed to be IID samples from a lognormal distribution, the we would expect an estimate for $\mathbb{E}\qty[f(s)]$ that is asymmetric across the $f(s)$ in different SOW, not an unweighted average.
}

\AR{
    This is another insightful comment, and I think we can frame this a bit better in the literature on non-naive Monte Carlo sampling.
    As noted, we write in section 2.3 that
}
\begin{quote}
    If the \glspl{sow} are drawn \gls{iid} from the true distribution, then the expected value of some function  $f$, $\mathbb{E}\qty[f(\vb{s})]$, can be readily estimated as $\frac{1}{N} \sum_{j=1}^N f(\vb{s}_j)$.
    However, this is often not the case; for example, a fixed set of simulations may be available for analysis or low-probability but high-impact regions of the parameter space may have been sampled.
    In this case, the formula must be adjusted to
    \begin{equation*}
        \mathbb{E}\qty[f(\vb{s})] \approx \frac{1}{N} \sum_{i=j}^N w_j f(\vb{s}_j),
    \end{equation*}
    where the $w_j$ are suitably chosen weights such that their sum is equal to 1.
\end{quote}

\AR{
    This does not require assuming that the set of sows, $\vb{s}$, is a discrete uniform RV, only that they are drawn IID from $p(\vb{s})$
}

\begin{quote}
    Need to revisit -- the comment about population doesn't make sense to me
\end{quote}

@@

\RC{
    I am further confused by equation 1 and the following paragraph on choosing the weights.
    First, are the weights “chosen” by the analyst, as stated in line 266, or are they estimated as a way of making clear what weights are implied by a certain selection of SOW?
}

\AR{
    Let's make sure it is clear that they are implied by a certain $p_\text{belief}$ and sampling distribution of SOWs
}

\begin{quote}
    Let's change it so that instead of saying ``suitably chosen'' weights we just say $\text{s.t.}~ \sum_j w_j = 1$.
    There's another set of questions.
\end{quote}

@@

\AR{
    Thanks, we can add a bit to the discussion.
}

\RC{
    Similarly, Figure 8 shows the key results from the weighting step.
    As I understand it, panel a shows the authors illustrative assumptions for sea level rise distributions, and then panels b-d show what weights across SOWs are consistent with those assumptions. What should a decision maker takeaway from this? How might they use it in their decision process? It would be helpful to see this articulated in the example and expanded on in the discussion or conclusions.
}

\AR{
    This should come across more clearly in the conclusions.
}

% ugly hack to hide month in APACite
\renewcommand{\APACrefYearMonthDay}[3]{\APACrefYear{#1}}
\bibliographystyle{apacite}
\bibliography{library-bibtex}

### 2022-09-30 papers/earths-future/submission_02.tex
ready for klaus comments on intro

\begin{keypoints}
  \item Decisions about how to manage climate risks are often made under deep uncertainty.
  \item We introduce a Bayesian framework to transparently synthesize and characterize deep uncertainties with the goal to support decision-making.
  \item We demonstrate the framework using a simple case study of house elevation for coastal flood risk management.
\end{keypoints}

\begin{abstract}
  Projections of future climate risks can vary considerably from one source to another, posing considerable communication and decision-analytical challenges.
  One such challenge is how to present trade-offs under deep uncertainty in a salient and interpretable manner.
  Some common approaches include analyzing a small subset of projections or invoking Laplace's principle of insufficient reason to justify a simple average.
  These approaches can underestimate risks, hide deep uncertainties, and provide little insight into which assumptions drive decision-relevant outcomes.
  Here we introduce and demonstrate a transparent Bayesian framework for synthesizing deep uncertainties to inform climate risk management.
  The first step of this workflow is to generate an ensemble of simulations representing possible futures and analyze them through standard exploratory modeling techniques.
  Next, a small set of probability distributions representing subjective beliefs about the likelihood of possible futures is used to weight the scenarios.
  Finally, these weights are used to compute and characterize trade-offs, conduct robustness checks, and reveal implicit assumptions.
  We demonstrate the framework through a didactic case study analyzing how high to elevate a house to manage coastal flood risks.
\end{abstract}

@@

The identification of sound strategies to manage risks driven by climatic changes is challenged by large uncertainties surrounding projections of the coupled natural-human systems that influence projections of climate risk. These uncertainties often arise from choices experts have to make, for example about how to formulate scientific models of future water levels. Different experts can disagree about these choices, leading to different projections. Analyzing decisions in such a situation of deep uncertainty poses nontrivial challenges. For example, picking a single representative projection can lead to under-estimation of risk and poor decisions. Similarly, communicating results separately for each projection can overwhelm decision-makers. To make matters worse, typical approaches to this problem are mostly silent on what assumptions make a difference for the decisions at hand. We develop and demonstrate a framework to address these challenges. The framework provides a transparent approach to (i) combine a large number of deeply uncertain projections to a more understandable and smaller sample set and (ii) provide insights about which assumptions and modeling choices influence decisions. We demonstrate the approach with a relatively simple example question of how high to elevate a house in the face of deeply uncertain projections of future water levels.

\section{Introduction}\label{sec:introduction}

Many critical infrastructure services around the world are aging \cite<\eg,>{ho_dams:2017}.
Moreover, changes in regulations, finance, patterns of population and infrastructure use, and the climate system challenge the ability of critical infrastructures to meet design objectives \cite{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020}.
The need to upgrade and expand infrastructure motivates the question: for which possible futures should infrastructure systems and components be designed?
The answer to this question depends in part on values: intrinsic trade-offs between safety, performance, cost, and other objectives depend on context and stakeholder preferences \cite{keller_management:2021} and are often regulated by statute or industry guidelines \cite{bruneau_multihazard:2017}.
For example, hospitals and critical infrastructure are generally designed to a higher standard of risk protection than ordinary buildings \cite{asce_7-10:2013}.
At the same time, performance depends upon future conditions, and so decisions about which scenario to design for are necessarily subject to implicit or explicit assumptions about the likelihood or possibility of different futures.

\subsection{Current practice}

Current practice in engineering, infrastructure design, and local governance relies heavily on standards that specify particular design events or conditions that buildings and infrastructure should safely withstand \cite{asce_7-10:2013,bruneau_multihazard:2017}.
These are often, though not always, informed by probabilistic analysis of relevant data.
For example, \gls{fema}, local governments, and engineering consultants produce local floodplain maps in many communities.
These trigger specific floodplain regulations, such as a requirement for homes in the 100-year floodplain with federally backed mortgages to be covered by flood insurance \cite{kousky_voucher:2014}.
Additionally, local building codes (based on guidance such as \citeA{asce_24-05:2006} or \citeA{FEMA_p-55:2011}) may require new construction in flood zones to be elevated freeboard above a nominal \gls{bfe}.
Probabilistic analysis also informs the design of large-scale infrastructure.
For example, some levees in the Netherlands are designed for a nominal annual failure probability of $\nicefrac{1}{4000}$ \cite{eijgenraam_flooding:2014}, while a seawall proposed as part of a \$29 billion coastal protection project for  Galveston Bay was designed by setting the nominal annual probability of overtopping to 1\% \cite[Appendix D.,~p.~2-59]{USACE_coastal:2021}.

There are many advantages to standards-based design.
In particular, these heuristics are scalable and explainable, they reduce the complexity of design analysis, and they are fair in at least a procedural sense.
However, reliance on these heuristics also faces limitations.
In particular, this sort of one-size-fits-all guidance may not be an efficient or desirable way to balance trade-offs between metrics of stakeholder values such as sense of place, distributive justice, economic efficiency, and safety \cite{keller_management:2021}.
This has motivated economically informed approaches, like risk-based design and cost benefit analysis \cite{eijgenraam_flooding:2014,vandantzig_dike:1956,xian_elevation:2017}, that place ``a strong emphasis upon a proportionate response to risk, so that the amount invested in risk reduction is in proportion to the magnitude of the risk and the cost-effectiveness with which that risk may be reduced'' \cite{merz_fluvial:2010}.
These quantitative cost-benefit analyses also rely on probabilistic descriptions of relevant hazard, and can be particularly sensitive to representation of tail probabilities \cite{merz_heavytails:2022,wong_floodrisk:2018,garner_slrise:2018}.

The approaches currently used to estimate the \glspl{pdf} of climate hazards used in standard-setting and decision-making emphasize nominally objective methods that can be applied consistently across locations.
For example, \gls{usgs} Bulletin 17C specifies procedures for estimating flood frequency \cite{bulletin17c:2019} and \gls{noaa} Atlas 14 provides estimates of the intensity, duration, and frequency of extreme rainfall \cite{atlas14_texas:2018}.
Among several statistical assumptions these analyses make that have been recently called into question is stationarity \cite<the assumption that past and future hazard come from the same \gls{pdf}; see>[for a review]{merz_review:2014}.
For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \cite{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}, and trends in many other hazards for other locations are consistent with observations \cite<see>[for a comprehensive summary]{ipcc_impacts:2022}.
While some methods have been proposed for incorporating trends into these analyses \cite<see>[for a review]{Salas:2018ge}, these assume specific forms of nonstationarity which may not fully represent physical processes or true levels of uncertainty \cite{doss-gollin_robustadaptation:2019,Montanari:2014hl,Serinaldi:2015bq}.
Ultimately, the difficulty of incorporating nonstationarity into existing frameworks has led to continued reliance on stationarity despite recognition of the associated limitations as discussed in p.~2 of \citeA{bulletin17c:2019} and p.~A.4-42 of \citeA{atlas14_texas:2018}.

\subsection{Emerging paradigms}

Projecting nonstationary hazard is difficult because many future hazards depend on human decisions (\eg, future greenhouse gas emissions) or on physical processes that are poorly constrained by existing data \cite<\eg, collapse of the West Antarctic ice sheet;>[]{deconto_antarctica:2016}.
In other words, they are deeply uncertain \cite{keller_management:2021,walker_deep:2013,lempert_complex:2002,haasnoot_sealevelrise:2021}.
These deeply uncertain nonstationary hazards challenge not only the existing stationary estimates but, more fundamentally, the premise that objective estimates of future hazard exist and can be estimated empirically.

Recognizing the challenges of \gls{dmdu}, many frameworks for identifying robust decisions have been proposed.
Most emphasize the use of models in an exploratory (``what-if'') framework to learn about interactions between decisions and system dynamics \cite{bankes:1993}.
For example, robust decision making \cite{lempert_shaping:2003} evaluates models over large ensembles of possible futures to assess the performance of different policies under each, then applies statistical analysis to identify the conditions under which particular policies perform well or poorly.
When the decision space is complex, many analyses use policy search (often using multi-objective optimization tools) to identify promising actions \cite{kasprzyk:2013,kasprzyk_denovo:2012,hadka_mordm:2015}.
In addition, many studies formally quantify robustness to deep uncertainties \cite{herman:2015,mcphail_robustness:2019} and use this as a criterion for policy comparison.

Although these \gls{dmdu} methods have proven valuable in a wide range of settings, they still require necessarily subjective assessments about the likelihood of future conditions.
Even exploratory models require the analyst to choose which uncertainties are considered and how they are sampled.
When analyses use metrics that integrate performance over many possible futures, whether to compute sensitivities, estimate expectations, or compute robustness metrics, they necessarily make assumptions about the likelihood of different futures.
For example, uniform priors are often justified by deferring to Laplace's principle of insufficient reason \cite<see>[p.~135]{stigler_uncertainty:1986}, but even this choice makes strong assumptions that are sensitive to factors such as parameterization \cite[p.54]{gelman_bda3:2014}.
While one interpretation of deep uncertainty is that that specifying a joint \gls{pdf} over inputs is inappropriate, assumptions about the ranges and independence of parameters to sample are just as subjective as the choice of probability distribution \cite{schneider_scenarios:2002,quinn_exploratory:2020} and can often be interpreted  as a specific choice of probability distribution (we revisit this point in \cref{sec:analysis-condition}).
This motivates the development of decision analytic frameworks that draw from the strengths of \gls{dmdu} methods such as exploratory modeling, vulnerability assessment, robustness checks, and iterative stakeholder critique, but that embrace the reality that assumptions about the future are inescapable.

\subsection{Research gaps and objectives}

We are motivated by parallels between decision making under deep uncertainty and the statistical problem of model selection.
\citeA{oreskes_verification:1994} argues that because natural systems are never closed and model results are never unique, validation and verification of models representing these systems is necessarily qualitative and subjective.
In the literature on decision making under deep uncertainty, this has led to recognition of the need to develop strategies that are robust to model imperfections.
There are parallels to this challenge in the Bayesian literature on model selection, particularly in the ``$\mathcal{M}$-open'' case in which there is no ``true'' model to identify and all models are imperfect representations of reality.
This is a theoretical contrast to the ``$\mathcal{M}$-closed'' case in which the data are assumed to come from a ``true'' model which is a member of the finite set of candidate models available to the analyst \cite{bernardo_bayesian:1994}.
Commonly used classical methods for model selection implicitly assume $\mathcal{M}$-closed case and are defined by their ability to asymptotically identify the true model from a set of candidate models.
By contrast, model selection in the $\mathcal{M}$-open case emphasizes a subjective view of the modeling process, for which probability offers a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \citeA{gelman_philosophy:2013}, \citeA{ramsey_probability:2016}, or \citeA{jaynes_probability:2003} for a discussion of Bayesian philosophy and \citeA{Piironen:2017eh} for a technical discussion of methods for model selection).
From a practical perspective, model selection in the $\mathcal{M}$-open case emphasizes building models iteratively, simulating the consequences of these models, and subsequently using these simulations to critique and improve them until an acceptable model (in the analyst's subjective judgement) is reached \cite{gelman_workflow:2020}.
Thus, where classical statistics, particularly as applied to environmental systems, has often emphasized objectivity, we draw from this Bayesian literature that  emphasizes a focus on building transparency in the presentation, analysis, and critique of models and results.

In this paper we offer a first conceptual step towards bridging the fields of \gls{dmdu} and Bayesian model building.
We consider a didactic case study of whether to elevate a hypothetical house, and if so how high, as a specific example of a decision problem subject to both shallow (storm surge) and deep (\gls{slr}) uncertainties.
Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \cite{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,rozer_coping:2016,mobley_mitigation:2020,aerts_cost:2018}, and both local building codes \cite{asce_7-10:2013,bruneau_multihazard:2017,asce_24-05:2006} and federal policy \cite{FEMA_p-55:2011} require elevation in some cases.
Guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \cite{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent research has demonstrated that neglecting uncertainty in the cost-benefit analysis can lead to poor decisions \cite{zarekarizi_suboptimal:2020}.
Focusing on deep uncertainty in \gls{slr} over the 70 year design life of a hypothetical house, we seek to answer the research question ``\emph{how can deep uncertainties be transparently synthesized for decision analysis?}''

>>>>

\begin{keypoints}
  \item Decisions about how to manage climate risks are often made under deep uncertainty.
  \item We introduce a Bayesian framework to transparently synthesize and characterize deep uncertainties with the goal to support decision-making.
  \item We demonstrate the framework using a simple case study of house elevation for coastal flood risk management.
  \item Estimates of performance or robustness under deep uncertainty necessarily involve subjective judgement.
\end{keypoints}

\begin{abstract}
  Projections of nonstationary climate risks can vary considerably from one source to another, posing considerable communication and decision-analytical challenges.
  One such challenge is how to present trade-offs under deep uncertainty in a salient and interpretable manner.
  Some common approaches include analyzing a small subset of projections or invoking treating all projections as equally likely regardless of how they are sampled.
  These approaches can underestimate risks, hide deep uncertainties, and provide little insight into which assumptions drive decision-relevant outcomes.
  Here we introduce and demonstrate a transparent Bayesian framework for synthesizing deep uncertainties to inform climate risk management.
  The first step of this workflow is to generate an ensemble of simulations representing possible futures and analyze them through standard exploratory modeling techniques.
  Next, a small set of probability distributions representing subjective beliefs about the likelihood of possible futures is used to weight the scenarios.
  Finally, these weights are used to compute and characterize trade-offs, conduct robustness checks, and reveal implicit assumptions.
  We demonstrate the framework through a didactic case study analyzing how high to elevate a house to manage coastal flood risks.
\end{abstract}

@@

The identification of sound strategies to manage risks driven by climatic changes is challenged by large uncertainties surrounding projections of the coupled natural-human systems that influence projections of climate risk.
These uncertainties often arise from choices experts have to make, for example about how to formulate scientific models of future water levels.
Different experts can disagree about these choices, leading to different projections.
Analyzing decisions in such a situation of deep uncertainty poses nontrivial challenges.
For example, picking a single representative projection can lead to under-estimation of risk and poor decisions.
Similarly, communicating results separately for each projection can overwhelm decision-makers.
To make matters worse, typical approaches to this problem are mostly silent on what assumptions make a difference for the decisions at hand.
We develop and demonstrate a framework to address these challenges.
The framework provides a transparent approach to (i) combine a large number of deeply uncertain projections to a more interpretable sample set and (ii) provide insights about which assumptions and modeling choices influence decisions.
We demonstrate the approach with a relatively simple example question of how high to elevate a house in the face of deeply uncertain projections of future water levels.

\clearpage
\section{Introduction}\label{sec:introduction}

Aging infrastructure and changes in regulations, finance, patterns of population and infrastructure use, and climate challenge the ability of critical infrastructures to meet design objectives \cite{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020,asce_infrastructure_climate:2021,ho_dams:2017}.
To ensure acceptable performance, current practice in engineering, infrastructure design, and regulation relies heavily on standards that specify design events or conditions that buildings and infrastructure should safely withstand \cite{bruneau_multihazard:2017}.
For example, \gls{fema}, local governments, and engineering consultants produce local floodplain maps in many communities.
Buildings in the designated floodplain are subject to specific regulations, such as flood insurance requirements as an eligibility requirement for federally backed mortgages \cite{kousky_voucher:2014} or minimum elevations \cite{asce_24-05:2006,FEMA_p-55:2011}.
Although this paper focuses on flooding, similar approaches inform mitigation strategies for a wide range of other hazards \cite{asce_7-10:2013}.

Standards-based risk management frameworks have many advantages, including scalability, explainability, and simplicity.
However, the choice of standard is a complex design and policy choice.
Risk-based design and cost benefit analysis \cite{eijgenraam_flooding:2014,vandantzig_dike:1956,xian_elevation:2017} can provide a quantitative basis for comparing possible standards by emphasizing ``a proportionate response to risk, so that the amount invested in risk reduction is in proportion to the magnitude of the risk and the cost-effectiveness with which that risk may be reduced'' \cite{merz_fluvial:2010}.
This provides a formal basis for, \eg, protecting hospitals and critical infrastructure to a higher degree than ordinary buildings \cite{asce_7-10:2013}.
However, these methods are silent on how standards should balance trade-offs, not only between cost and performance but also between other stakeholder values such as sense of place, distributive justice, and safety \cite{keller_management:2021}.

Moreover, estimates of the performance trade-offs require implicit or explicit assumptions about the likelihood of different possible futures.
Current practice emphasizes nominally objective methods that can be applied consistently across locations.
For example, the \gls{usgs} Bulletin 17C specifies procedures for estimating flood frequency \cite{bulletin17c:2019}.
Similarly, \gls{noaa} Atlas 14 provides estimates of the intensity, duration, and frequency of extreme rainfall \cite{atlas14_texas:2018,noaa_nonstationary:2022}.
Among several statistical assumptions these analyses make that have been recently called into question is stationarity \cite<the assumption that past and future hazard come from the same \gls{pdf}>{merz_review:2014,milly_stationarity:2008,doss-gollin_robustadaptation:2019}.
While some methods have been proposed for incorporating nonstationarity into risk analyses \cite<see>[for a review]{Salas:2018ge}, these assume specific forms of a trend which may not adequately represent physical processes or true levels of uncertainty \cite{doss-gollin_robustadaptation:2019,montanari_immortal:2014,serinaldi_undead:2015}.
At least in part because of the challenges associated with developing objective methods to select from diverging projections of future hazard, official guidance continues to rely on the stationarity assumption \cite{bulletin17c:2019,atlas14_texas:2018}.

The limitations of objectivist approaches to projecting risk extend beyond estimating nonstationary climate hazards.
Human-natural systems are never closed and model results are never unique, and thus validation and verification of models representing these systems is necessarily qualitative and subjective \cite{oreskes_verification:1994}.
In other words, no model exists that could represent the full truth, and the future is thus deeply uncertain \cite{keller_management:2021,walker_deep:2013,lempert_complex:2002,haasnoot_sealevelrise:2021}.
Consequently, a growing literature on \gls{dmdu} emphasizes the value of identifying decisions that are robust, in some sense, to model limitations \cite{moody_robustness:2013,herman:2015,mcphail_robustness:2019,borgomeo_robustness:2018}.
Within this literature has emerged a debate regarding the value and use of probabilistic information  \cite<see>[and references therein]{taner_probabilistic:2019}.
On the one hand, scholars have pointed out that predictions are inherently unreliable, and representing deep uncertainties through probability distributions frequently over-estimates predictive skill \cite{groves_scenarios:2007,lempert_robust:2000}.
On the other, assessments of which decisions are robust depend on subjective choices about how to define robustness and how to sample uncertainties \cite{mcphail_robustness:2019,quinn_exploratory:2020,schneider_scenarios:2002,schneider_dangerous:2001,reis_distribution:2020}.

In this paper we offer a first conceptual step towards bridging this divide by presenting a framework that combines the strengths of both approaches.
In the first step, exploratory or bottom-up modeling is used to build insight and identify potential system vulnerabilities \cite{moallemi_exploratory:2020,bankes:1993,brown_decisionscaling:2012}.
In the second, we synthesize the results of exploratory modeling to formally estimate performance metrics and trade-offs using probabilistic models for uncertainty.
Drawing from the literature on building predictive models when ``all models are wrong'' \cite{box_sciencestatistics:1976,gelman_philosophy:2013,Piironen:2017eh}, we interpret these probability distributions not statements of fact, but rather as a self-consistent framework for reasoning about how different assumptions lead to different inferences.
An advantage of our approach is that it facilitates computationally efficient analysis of how alternative probabilistic models would affect estimated performance metrics and trade-offs.

We illustrate our approach through a didactic case study of whether to elevate a hypothetical house, and if so how high.
Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \cite{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,rozer_coping:2016,mobley_mitigation:2020,aerts_cost:2018}, and both local building codes \cite{asce_7-10:2013,bruneau_multihazard:2017,asce_24-05:2006} and federal policy \cite{FEMA_p-55:2011} require elevation in some cases.
Guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \cite{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent research has demonstrated that neglecting uncertainty in the cost-benefit analysis can lead to poor decisions \cite{zarekarizi_suboptimal:2020}.
Focusing on deep uncertainty in \gls{slr} over the 70 year design life of a hypothetical house, we seek to answer the research question ``\emph{how can deep uncertainties be transparently synthesized for decision analysis?}''
To shed light on this question in a single paper, we necessarily are silent on key issues that would be relevant to real-world decision-makers including alternative decision levers, the potential for adaptive decision rules, and rival problem framings.

We proceed as follows.
In \cref{sec:analysis} we present three formal decision analytic frameworks for analyzing an ensemble of \gls{slr} simulations, building through existing approaches for exploratory modeling (\cref{sec:analysis-explore}) and scenario analysis (\cref{sec:analysis-condition}) towards a formal Bayesian method for transparently synthesizing deep uncertainty through subjective prior beliefs (\cref{sec:analysis-synthesize}).
In \cref{sec:case-study} we describe the case study in detail.
Next, in \cref{sec:results}, we present results for each of the three decision lenses and discuss the advantages and limitations of each theoretical approach.
In \cref{sec:caveats} we discuss limitations of the study and future research needs.
Finally in \cref{sec:conclusions} we discuss key findings and implications for policy and practice.
